---
title: "230A - Group Project"
output: pdf_document
date: "2024-04-29"
---

We downloaded the data from here: https://www2.census.gov/programs-surveys/acs/data/pums/2022/1-Year/csv_pus.zip

First, we append the files and then we process the data to reduce the size and 
generate some variables we need
```{r}
if (!require("readr")) install.packages("readr", dependencies = TRUE)
library(readr)

data_a <- read.csv("/Users/zequi/Desktop/Berkeley/230A/csv_pus/psam_pusa.csv")
data_b <- read.csv("/Users/zequi/Desktop/Berkeley/230A/csv_pus/psam_pusb.csv")
data <- rbind(data_a, data_b)

if (!require("dplyr")) install.packages("dplyr", dependencies = TRUE)
library(dplyr)

data <- data %>%
  select(ST, DIVISION, AGEP, ESR, CIT, NATIVITY, CITWP, YOEP, COW, DIS, ENG, PAP, FER, MAR, SEX, SCH, SCHL, PINCP, WAGP, ANC, ANC1P, ANC2P, ESR, MIGSP, POBP, RAC1P, PERNP, HISP, HINS4, ADJINC)

write.csv(data, "psam_pus.csv", row.names = FALSE)
zip("psam_pus.zip", "psam_pus.csv")
file.remove("psam_pus.csv")

processed_data <- data %>%
  filter(AGEP >= 18, !ESR %in% c("b", 6)) %>%
  mutate(
    Citizen = ifelse(CIT %in% c(1, 2, 3, 4), 1, 0),
    Foreign_born = ifelse(NATIVITY == 2, 1, 0),
    High_School_Grad = ifelse(SCHL >= 16 & SCHL < 21, 1, 0),
    Bachelors_Degree = ifelse(SCHL >= 21, 1, 0),
    Unemployed = ifelse(ESR == 3, 1, 0),
    HISP_dummy = ifelse(HISP == 1, 0, 1),
    Years_in_US = 2022 - YOEP,
    Years_naturalized = 2022 - CITWP,
    SEX = as.numeric(SEX) - 1,
    HINS4 = as.numeric(HINS4) - 1,
    Asian = ifelse(RAC1P ==6, 1, 0)
  )

write.csv(processed_data, "processed_data.csv", row.names = FALSE)
zip("processed_data.zip", "processed_data.csv")
file.remove("processed_data.csv")
```

We bring the processed data:
```{r}
zip_file_path <- "/Users/zequi/Desktop/Berkeley/230A/stat230A/processed_data.zip"
csv_file_inside_zip <- "processed_data.csv"

processed_data <- read.csv(unzip(zip_file_path, files = csv_file_inside_zip))
```

We do some extra processing:
```{r}
library(tidyr)
library(dplyr)

processed_data <- processed_data %>%
  mutate(
    DIS = ifelse(DIS == 2, 0, 1), 
    HINS4 = 1 - as.numeric(HINS4),
    Citizen = ifelse(CIT %in% c(1, 3, 4), 1, 0),
    High_School_Grad = ifelse(SCHL >= 16, 1, 0)
  )

processed_data$ENG[is.na(processed_data$ENG)] <- 0

processed_data$wb50birth <- processed_data$FER
processed_data$wb50birth[is.na(processed_data$FER) &
                         (processed_data$AGEP >= 50 | processed_data$SEX == 0)] <- 0
processed_data = processed_data%>%filter(CIT != 2)
processed_data$Years_naturalized[processed_data$Foreign_born == 0] <- 0
processed_data$Years_in_US[processed_data$Foreign_born == 0] <- 0
processed_data$Years_in_US[processed_data$CIT == 5] <- 0
processed_data$adj_WAGP = processed_data$WAGP + 1
processed_data = processed_data%>%mutate(immigrant = CIT%in%c(4,5))
```

```{r}
processed_data2 = processed_data%>%filter(!(WAGP == 0 & Unemployed == 0))

processed_data2 = processed_data2%>%mutate(immigrant = as.integer(CIT%in%c(4,5)))

processed_data2 <- processed_data2%>%mutate(white_dummy = (RAC1P == 1),
                         asian_dummy = (RAC1P == 6),
                         black_dummy = (RAC1P == 2))
```


```{r}
immigrants_only = processed_data2%>%filter(CIT%in% c(4,5))
immigrants_only2 = immigrants_only%>%filter(RAC1P%in%c(1,2,6,8,9))
non_immigrants = processed_data2%>%filter(immigrant==1, RAC1P%in%c(1,2,6,8,9))
```

We will use this later to calculate robust se:
```{r}
library(gridExtra)
library(MASS)
library(sandwich)
library(knitr)

compute_ehw_se <- function(model) {
  robust_se <- sqrt(diag(vcovHC(model, type = "HC1")))
  return(robust_se)
}
```

Logistics regression: Predict Employment status
```{r}
employment_status_model <- glm(Unemployed ~  as.factor(RAC1P) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + as.factor(MAR) , data = processed_data2, family = binomial())

summary_model <- summary(employment_status_model)
```

```{r}
model <- employment_status_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
library(ggplot2)
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Unemployment-General",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Logistics regression: Risk of poverty
```{r}
processed_data2$PAP_binary = as.numeric(processed_data2$PAP > 0)
poverty_risk_model <- glm(PAP_binary ~ as.factor(RAC1P) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + as.factor(MAR), family = binomial(), data = processed_data2)
summary_model <- summary(poverty_risk_model)
```

```{r}
model <- poverty_risk_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Poverty risk (Public assistance income >0)-General",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Logistics regression: Probability of being on medicaid (HINS4)
```{r}
medicaid_model <- glm(HINS4 ~ as.factor(RAC1P) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + as.factor(MAR), family = binomial(), data = processed_data2)

summary_model <- summary(medicaid_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]
summary_model
```

```{r}
model <- medicaid_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Medicaid coverage-General",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Repeat the exercise for leading Hispanic origins
```{r}
hisp_only = processed_data2%>%filter(HISP_dummy==1)
hisp_only_im = immigrants_only2%>%filter(HISP_dummy==1)
```

```{r}
largest_subgroup = hisp_only_im%>%group_by(POBP)%>%
  summarise(count = n())%>%
  mutate(per = (count/sum(count))*100)%>%
  arrange(desc(per))%>%head(5)%>%pull(POBP)
```

```{r}
hisp_only_im = hisp_only_im%>%filter(POBP %in% largest_subgroup)
```

Hispanic immigrants
```{r}
imm_hisp_only_employment_status_model <- glm(Unemployed ~  as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), data = hisp_only_im, family = binomial())
summary_model <- summary(imm_hisp_only_employment_status_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_hisp_only_employment_status_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Unemployment-Hispanic immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
hisp_only_im$PAP_binary = as.numeric(hisp_only_im$PAP > 0)
imm_hisp_only_poverty_risk_model <- glm(PAP_binary ~ as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = hisp_only_im)
summary_model <- summary(imm_hisp_only_poverty_risk_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_hisp_only_poverty_risk_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Poverty risk-Hispanic immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
imm_hisp_only_medicaid_model <- glm(HINS4 ~ as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = hisp_only_im)
summary_model <- summary(imm_hisp_only_medicaid_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_hisp_only_medicaid_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Medicaid coverage-Hispanic immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Hispanics immigrants and native-born:
```{r}
hisp_only_employment_status_model <- glm(Unemployed ~ AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), data = hisp_only, family = binomial())
summary_model <- summary(hisp_only_employment_status_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- hisp_only_employment_status_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Unemployment-Hispanic immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
hisp_only_poverty_risk_model <- glm(PAP_binary ~ AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = hisp_only)
summary_model <- summary(hisp_only_poverty_risk_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- hisp_only_poverty_risk_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Poverty risk-Hispanic immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
hisp_only_medicaid_model <- glm(HINS4 ~ AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = hisp_only)
summary_model <- summary(hisp_only_medicaid_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- hisp_only_medicaid_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Medicaid coverage-Hispanic immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Repeat the exercise for leading Asian origins
```{r}
asian_only = processed_data2%>%filter(RAC1P==6)
asian_only_im = immigrants_only2%>%filter(RAC1P==6)
```

```{r}
largest_subgroup = asian_only_im%>%group_by(POBP)%>%
  summarise(count = n())%>%
  mutate(per = (count/sum(count))*100)%>%
  arrange(desc(per))%>%head(5)%>%pull(POBP)
```

```{r}
asian_only_im = asian_only_im%>%filter(POBP %in% largest_subgroup)
```

Asian immigrants
```{r}
imm_asian_only_employment_status_model <- glm(Unemployed ~  as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), data = asian_only_im, family = binomial())
summary_model <- summary(imm_asian_only_employment_status_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_asian_only_employment_status_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```

```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Unemployment-Asian immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
asian_only_im$PAP_binary = as.numeric(asian_only_im$PAP > 0)
imm_asian_only_poverty_risk_model <- glm(PAP_binary ~ as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = asian_only_im)
summary_model <- summary(imm_asian_only_poverty_risk_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_asian_only_poverty_risk_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Poverty risk-Asian immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
imm_asian_only_medicaid_model <- glm(HINS4 ~ as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = asian_only_im)
summary_model <- summary(imm_asian_only_medicaid_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_asian_only_medicaid_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Medicaid coverage-Asian immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Asian immigrants and native
```{r}
asian_only_employment_status_model <- glm(Unemployed ~ AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), data = asian_only, family = binomial())
summary_model <- summary(asian_only_employment_status_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- asian_only_employment_status_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Unemployment-Asian immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
asian_only_poverty_risk_model <- glm(PAP_binary ~  AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = asian_only)
summary_model <- summary(asian_only_poverty_risk_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- asian_only_poverty_risk_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Poverty risk-Asian immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
asian_only_medicaid_model <- glm(HINS4 ~ AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = asian_only)
summary_model <- summary(asian_only_medicaid_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- asian_only_medicaid_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

coeff_df <- as.data.frame(filtered_coefficients)

coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

print(coeff_df, digits=3)
```
```{r}
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Medicaid coverage-Asian immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

EDA
```{r}
library(ggplot2)
library(gridExtra)
library(scales)

generate_and_save_plots <- function(data, variable, file_name) {
  base_plot <- ggplot(data, aes_string(x = variable, y = "..prop..", group = "1")) +
    geom_bar(aes(y = ..prop..), stat = "count", fill = "gray") +
    scale_y_continuous(labels = percent_format()) +
    geom_text(stat = "count", aes(label = percent(..prop..), y = ..prop..), 
              vjust = -0.5, position = position_stack(vjust = 0.5))

  p1 <- base_plot +
    labs(title = paste("General Distribution of", variable), x = variable, y = "Proportion")
  
  p2 <- base_plot +
    facet_grid(. ~ Citizen) +
    labs(title = paste("Distribution of", variable, "by Citizen Status"), x = variable, y = "Proportion")
  
  p3 <- base_plot +
    facet_grid(. ~ immigrant) +
    labs(title = paste("Distribution of", variable, "by Immigrant Status"), x = variable, y = "Proportion")

  p4 <- base_plot +
    facet_grid(. ~ asian_dummy) +
    labs(title = paste("Distribution of", variable, "by Asian identification"), x = variable, y = "Proportion")

  p5 <- base_plot +
    facet_grid(. ~ HISP_dummy) +
    labs(title = paste("Distribution of", variable, "by Hispanic identification"), x = variable, y = "Proportion")
  
  combined_plot <- grid.arrange(p1, p2, p3, p4, p5, ncol = 1)
  
  ggsave(file_name, combined_plot, width = 8, height = 30)  # Adjusted height to accommodate more plots
}

generate_and_save_plots(processed_data2, "PAP_binary", "plots_PAP_binary.png")
generate_and_save_plots(processed_data2, "HINS4", "plots_HINS4.png")
generate_and_save_plots(processed_data2, "Unemployed", "plots_Unemployed.png")
```
