---
title: "Ezequiel_markdown"
output: pdf_document
date: "2024-04-29"
---

We downloaded the data from here: https://www2.census.gov/programs-surveys/acs/data/pums/2022/1-Year/csv_pus.zip

First, we append the files and then we process the data to reduce the size and 
generate some variables we need
```{r}
if (!require("readr")) install.packages("readr", dependencies = TRUE)
library(readr)

# We append population a, b, c,d  data once
data_a <- read.csv("/Users/zequi/Desktop/Berkeley/230A/csv_pus/psam_pusa.csv")
data_b <- read.csv("/Users/zequi/Desktop/Berkeley/230A/csv_pus/psam_pusb.csv")
data <- rbind(data_a, data_b)

if (!require("dplyr")) install.packages("dplyr", dependencies = TRUE)
library(dplyr)

data <- data %>%
  select(ST, DIVISION, AGEP, ESR, CIT, NATIVITY, CITWP, YOEP, COW, DIS, ENG, PAP, FER, MAR, SEX, SCH, SCHL, PINCP, WAGP, ANC, ANC1P, ANC2P, ESR, MIGSP, POBP, RAC1P, PERNP, HISP, HINS4, ADJINC)

write.csv(data, "psam_pus.csv", row.names = FALSE)
zip("psam_pus.zip", "psam_pus.csv")
file.remove("psam_pus.csv")

processed_data <- data %>%
  filter(AGEP >= 18, !ESR %in% c("b", 6)) %>%
  mutate(
    Citizen = ifelse(CIT %in% c(1, 2, 3, 4), 1, 0),
    Foreign_born = ifelse(NATIVITY == 2, 1, 0),
    High_School_Grad = ifelse(SCHL >= 16 & SCHL < 21, 1, 0),
    Bachelors_Degree = ifelse(SCHL >= 21, 1, 0),
    Unemployed = ifelse(ESR == 3, 1, 0),
    HISP_dummy = ifelse(HISP == 1, 0, 1),
    Years_in_US = 2022 - YOEP,
    Years_naturalized = 2022 - CITWP,
    SEX = as.numeric(SEX) - 1,
    HINS4 = as.numeric(HINS4) - 1,
    Asian = ifelse(RAC1P ==6, 1, 0)
  )

write.csv(processed_data, "processed_data.csv", row.names = FALSE)
zip("processed_data.zip", "processed_data.csv")
file.remove("processed_data.csv")
```

We bring the processed data:
```{r}
zip_file_path <- "/Users/zequi/Desktop/Berkeley/230A/stat230A/processed_data.zip"
csv_file_inside_zip <- "processed_data.csv"

processed_data <- read.csv(unzip(zip_file_path, files = csv_file_inside_zip))
```

```{r}
library(gridExtra)
library(MASS)
library(sandwich)
library(knitr)

compute_ehw_se <- function(model) {
  robust_se <- sqrt(diag(vcovHC(model, type = "HC1")))
  return(robust_se)
}
```

We do some extra processing:
```{r}
library(tidyr)
library(dplyr)

processed_data <- processed_data %>%
  mutate(
    DIS = ifelse(DIS == 2, 0, 1), 
    HINS4 = 1 - as.numeric(HINS4),
    Citizen = ifelse(CIT %in% c(1, 3, 4), 1, 0),
    High_School_Grad = ifelse(SCHL >= 16, 1, 0)
  )

processed_data$ENG[is.na(processed_data$ENG)] <- 0

processed_data$wb50birth <- processed_data$FER
processed_data$wb50birth[is.na(processed_data$FER) &
                         (processed_data$AGEP >= 50 | processed_data$SEX == 0)] <- 0
processed_data = processed_data%>%filter(CIT != 2)
processed_data$Years_naturalized[processed_data$Foreign_born == 0] <- 0
processed_data$Years_in_US[processed_data$Foreign_born == 0] <- 0
processed_data$Years_in_US[processed_data$CIT == 5] <- 0
processed_data$adj_WAGP = processed_data$WAGP + 1
processed_data = processed_data%>%mutate(immigrant = CIT%in%c(4,5))
```

```{r}
processed_data2 = processed_data%>%filter(!(WAGP == 0 & Unemployed == 0))
```

```{r}
processed_data2 = processed_data2%>%mutate(immigrant = as.integer(CIT%in%c(4,5)))
```

```{r}
processed_data2 <- processed_data2%>%mutate(white_dummy = (RAC1P == 1),
                         asian_dummy = (RAC1P == 6),
                         black_dummy = (RAC1P == 2))
```


```{r}
immigrants_only = processed_data2%>%filter(CIT%in% c(4,5))
immigrants_only2 = immigrants_only%>%filter(RAC1P%in%c(1,2,6,8,9))
```

```{r}
non_immigrants = processed_data2%>%filter(immigrant==1, RAC1P%in%c(1,2,6,8,9))
```

Logistics regression: Predict Employment status
```{r}
employment_status_model <- glm(Unemployed ~  as.factor(RAC1P) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + as.factor(MAR) , data = processed_data2, family = binomial())

summary_model <- summary(employment_status_model)
```

```{r}
model <- employment_status_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
library(ggplot2)
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Unemployment-General",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Logistics regression: Risk of poverty
```{r}
processed_data2$PAP_binary = as.numeric(processed_data2$PAP > 0)
poverty_risk_model <- glm(PAP_binary ~ as.factor(RAC1P) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + as.factor(MAR), family = binomial(), data = processed_data2)
#summary(poverty_risk_model)

summary_model <- summary(poverty_risk_model)
```

```{r}
model <- poverty_risk_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Poverty risk (Public assistance income >0)-General",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Logistics regression: Probability of being on medicaid (HINS4)
```{r}
medicaid_model <- glm(HINS4 ~ as.factor(RAC1P) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + as.factor(MAR), family = binomial(), data = processed_data2)
#summary(medicaid_model)

summary_model <- summary(medicaid_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]
summary_model
```

```{r}
model <- medicaid_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Medicaid coverage-General",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Repeat the exercise for top Hispanic origins
```{r}
hisp_only = processed_data2%>%filter(HISP_dummy==1)
hisp_only_im = immigrants_only2%>%filter(HISP_dummy==1)
```

```{r}
largest_subgroup = hisp_only_im%>%group_by(POBP)%>%
  summarise(count = n())%>%
  mutate(per = (count/sum(count))*100)%>%
  arrange(desc(per))%>%head(5)%>%pull(POBP)
```

```{r}
hisp_only_im = hisp_only_im%>%filter(POBP %in% largest_subgroup)
```

Hispanic immigrants

```{r}
imm_hisp_only_employment_status_model <- glm(Unemployed ~  as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), data = hisp_only_im, family = binomial())
summary_model <- summary(imm_hisp_only_employment_status_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_hisp_only_employment_status_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Unemployment-Hispanic immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
hisp_only_im$PAP_binary = as.numeric(hisp_only_im$PAP > 0)
imm_hisp_only_poverty_risk_model <- glm(PAP_binary ~ as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = hisp_only_im)
summary_model <- summary(imm_hisp_only_poverty_risk_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_hisp_only_poverty_risk_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Poverty risk-Hispanic immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
imm_hisp_only_medicaid_model <- glm(HINS4 ~ as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = hisp_only_im)
summary_model <- summary(imm_hisp_only_medicaid_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_hisp_only_medicaid_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Medicaid coverage-Hispanic immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Hispanics immigrants and native

```{r}
hisp_only_employment_status_model <- glm(Unemployed ~ AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), data = hisp_only, family = binomial())
summary_model <- summary(hisp_only_employment_status_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- hisp_only_employment_status_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Unemployment-Hispanic immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
hisp_only_poverty_risk_model <- glm(PAP_binary ~ AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = hisp_only)
summary_model <- summary(hisp_only_poverty_risk_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- hisp_only_poverty_risk_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Poverty risk-Hispanic immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
hisp_only_medicaid_model <- glm(HINS4 ~ AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = hisp_only)
summary_model <- summary(hisp_only_medicaid_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- hisp_only_medicaid_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Medicaid coverage-Hispanic immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Repeat the exercise for top Asian origins
```{r}
asian_only = processed_data2%>%filter(RAC1P==6)
asian_only_im = immigrants_only2%>%filter(RAC1P==6)
```

```{r}
largest_subgroup = asian_only_im%>%group_by(POBP)%>%
  summarise(count = n())%>%
  mutate(per = (count/sum(count))*100)%>%
  arrange(desc(per))%>%head(5)%>%pull(POBP)
```

```{r}
asian_only_im = asian_only_im%>%filter(POBP %in% largest_subgroup)
```

Asian immigrants
```{r}
imm_asian_only_employment_status_model <- glm(Unemployed ~  as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), data = asian_only_im, family = binomial())
summary_model <- summary(imm_asian_only_employment_status_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_asian_only_employment_status_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```

```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Unemployment-Asian immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```


```{r}
asian_only_im$PAP_binary = as.numeric(asian_only_im$PAP > 0)
imm_asian_only_poverty_risk_model <- glm(PAP_binary ~ as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = asian_only_im)
summary_model <- summary(imm_asian_only_poverty_risk_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_asian_only_poverty_risk_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Poverty risk-Asian immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```


```{r}
imm_asian_only_medicaid_model <- glm(HINS4 ~ as.factor(POBP) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = asian_only_im)
summary_model <- summary(imm_asian_only_medicaid_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- imm_asian_only_medicaid_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Medicaid coverage-Asian immigrants from five leading countries",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

Asian immigrants and native
```{r}
asian_only_employment_status_model <- glm(Unemployed ~ AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), data = asian_only, family = binomial())
summary_model <- summary(asian_only_employment_status_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- asian_only_employment_status_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Unemployment-Asian immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
asian_only_poverty_risk_model <- glm(PAP_binary ~  AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = asian_only)
summary_model <- summary(asian_only_poverty_risk_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- asian_only_poverty_risk_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Poverty risk-Asian immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```

```{r}
asian_only_medicaid_model <- glm(HINS4 ~ AGEP + SEX*immigrant + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR), family = binomial(), data = asian_only)
summary_model <- summary(asian_only_medicaid_model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]
summary_model$coefficients <- filtered_coefficients
print(summary_model$coefficients)
```

```{r}
model <- asian_only_medicaid_model
summary_model <- summary(model)
orig_coefficients <- coef(model)

ehw_se <- sqrt(diag(vcovHC(model, type = "HC0")))

# Get summary and filter out state coefficients
summary_model <- summary(model)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
filtered_coefficients <- summary_model$coefficients[!state_rows, ]

# Convert the matrix to a data frame for further manipulation
coeff_df <- as.data.frame(filtered_coefficients)

# Add robust z-values and p-values
coeff_df$Robust_z_value <- coeff_df$Estimate / ehw_se[!state_rows]
coeff_df$Robust_Pr <- 2 * pnorm(abs(coeff_df$Robust_z_value), lower.tail = FALSE)

# Assign significance stars based on the robust p-values
coeff_df$Signif_Stars <- ifelse(coeff_df$Robust_Pr < 0.001, "***", 
                                ifelse(coeff_df$Robust_Pr < 0.01, "**", 
                                       ifelse(coeff_df$Robust_Pr < 0.05, "*", 
                                              ifelse(coeff_df$Robust_Pr < 0.1, ".", ""))))

# Printing the formatted summary
print(coeff_df, digits=3)
```
```{r}
# Data frame for results
results <- data.frame(
  Coefficient = coeff_df$Estimate,
  SE = ehw_se[!state_rows],
  Significance = coeff_df$Signif_Stars
)
rownames(results) <- rownames(coeff_df)

# Prepare for plotting
results <- results[order(results$Coefficient), ]
results$Feature <- factor(rownames(results), levels = rownames(results))

# Plot
ggplot(results, aes(x = Feature, y = Coefficient)) +
  geom_point(aes(color = Significance), alpha = 0.5) +
  geom_errorbar(aes(ymin = Coefficient - SE, ymax = Coefficient + SE, color = Significance), width = 0.2) +
  coord_flip() +
  labs(title = "Coefficient Plot with EHW Robust Standard Errors",
       subtitle = "Medicaid coverage-Asian immigrants and native born",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() + theme(axis.text.y = element_text(size = 8))
```


Blinder Oxaca

```{r}
if (!require(oaxaca)) install.packages("oaxaca")
library(oaxaca)

oaxaca_data <- processed_data2[processed_data2$PERNP > 0, ]
oaxaca_data$PERNP <- as.numeric(oaxaca_data$PERNP)
oaxaca_data <- oaxaca_data[!(oaxaca_data$HISP_dummy == 1 & oaxaca_data$Asian == 1), ]
```

Ethnic Blinder Oxaca
```{r}
oaxaca_data$Ethnic_interest <- ifelse(oaxaca_data$HISP_dummy == 1 | oaxaca_data$Asian == 1, 1, 0)
oaxaca_data$Ethnic_interest <- as.factor(oaxaca_data$Ethnic_interest)

ethnic_oaxaca <- oaxaca(
  formula = log(adj_WAGP) ~ AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR) + PAP + HINS4 | Ethnic_interest,
  data = oaxaca_data
)
View(summary(ethnic_oaxaca))
```

```{r}
summary_model <- summary(ethnic_oaxaca$reg$reg.A)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]
print(summary_model$coefficients)
```

```{r}
summary_model <- summary(ethnic_oaxaca$reg$reg.B)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]
print(summary_model$coefficients)
```


```{r}
library(ggplot2)

overall_effects <- data.frame(
  Effect = c("Endowments", "Coefficients", "Interaction"),
  Coefficient = c(ethnic_oaxaca$threefold$overall["coef(endowments)"],
                  ethnic_oaxaca$threefold$overall["coef(coefficients)"],
                  ethnic_oaxaca$threefold$overall["coef(interaction)"]),
  StandardError = c(ethnic_oaxaca$threefold$overall["se(endowments)"],
                    ethnic_oaxaca$threefold$overall["se(coefficients)"],
                    ethnic_oaxaca$threefold$overall["se(interaction)"])
)

variable_effects <- data.frame(
  Variable = rownames(ethnic_oaxaca$threefold$variables),
  Endowments = ethnic_oaxaca$threefold$variables[, "coef(endowments)"],
  Endowments_SE = ethnic_oaxaca$threefold$variables[, "se(endowments)"],
  Coefficients = ethnic_oaxaca$threefold$variables[, "coef(coefficients)"],
  Coefficients_SE = ethnic_oaxaca$threefold$variables[, "se(coefficients)"]
)

selected_variables <- c("AGEP", "SEX", "Citizen", "as.factor(ENG)2", "as.factor(ENG)3",
                        "as.factor(ENG)4", "DIS", "Unemployed", "High_School_Grad",
                        "Bachelors_Degree", "PAP", "HINS4",
                        "SEX:Citizen", "SEX:DIS")

filtered_variable_effects <- variable_effects[variable_effects$Variable %in% selected_variables, ]

ggplot(filtered_variable_effects, aes(x = Variable, y = Endowments, fill = "Endowments")) +
  geom_col(color = "black", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = Endowments - Endowments_SE, ymax = Endowments + Endowments_SE), width = 0.2, position = position_dodge(width = 0.9)) +
  labs(title = "Endowment Effects for Selected Variables - Ethnic interest", x = NULL, y = "Endowment Effect") +
  scale_fill_manual(values = c("Endowments" = "white")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(filtered_variable_effects, aes(x = Variable, y = Coefficients, fill = "Coefficients")) +
  geom_col(color = "black", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = Coefficients - Coefficients_SE, ymax = Coefficients + Coefficients_SE), width = 0.2, position = position_dodge(width = 0.9)) +
  labs(title = "Coefficient Effects for Selected Variables - Ethnic interest", x = NULL, y = "Coefficient Effect") +
  scale_fill_manual(values = c("Coefficients" = "white")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
coefficients_A <- summary(ethnic_oaxaca$reg$reg.A)$coefficients
coefficients_A <- coefficients_A[!grepl("^as\\.factor\\(ST\\)|\\(Intercept\\)", rownames(coefficients_A)), ]
coefficients_A_df <- data.frame(
  Variable = rownames(coefficients_A),
  Coefficient = coefficients_A[, "Estimate"],
  StdError = coefficients_A[, "Std. Error"],
  Model = "Regression Asians/hispanics"
)

coefficients_B <- summary(ethnic_oaxaca$reg$reg.B)$coefficients
coefficients_B <- coefficients_B[!grepl("^as\\.factor\\(ST\\)|\\(Intercept\\)", rownames(coefficients_B)), ]
coefficients_B_df <- data.frame(
  Variable = rownames(coefficients_B),
  Coefficient = coefficients_B[, "Estimate"],
  StdError = coefficients_B[, "Std. Error"],
  Model = "Regression non-Asians and non-Hispanics"
)

all_coefficients <- rbind(coefficients_A_df, coefficients_B_df)

ggplot(all_coefficients, aes(x = Variable, y = Coefficient, fill = Model)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = Coefficient - StdError, ymax = Coefficient + StdError), 
                position = position_dodge(width = 0.8), width = 0.25) +
  coord_flip() +
  scale_fill_manual(values = c("grey70", "grey30")) +
  labs(title = "Comparison of Coefficients",
       x = "Variables",
       y = "Coefficients") +
  theme_minimal() +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold", hjust = 0.5, vjust = 1, size = 12),
        plot.margin = unit(c(1.5, 1, 1, 1), "lines"))  # Adjust the top margin to give more space for the title
```

Migrant Blinder Oxaca
```{r}
migrant_oaxaca <- oaxaca(
  formula = log(adj_WAGP) ~  AGEP + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + as.factor(MAR) + PAP + HINS4 | immigrant,
  data = oaxaca_data
)
View(summary(migrant_oaxaca))
```

```{r}
summary_model <- summary(migrant_oaxaca$reg$reg.A)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]
print(summary_model$coefficients)
```

```{r}
summary_model <- summary(migrant_oaxaca$reg$reg.B)
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]
print(summary_model$coefficients)
```

```{r}
overall_effects <- data.frame(
  Effect = c("Endowments", "Coefficients", "Interaction"),
  Coefficient = c(migrant_oaxaca$threefold$overall["coef(endowments)"],
                  migrant_oaxaca$threefold$overall["coef(coefficients)"],
                  migrant_oaxaca$threefold$overall["coef(interaction)"]),
  StandardError = c(migrant_oaxaca$threefold$overall["se(endowments)"],
                    migrant_oaxaca$threefold$overall["se(coefficients)"],
                    migrant_oaxaca$threefold$overall["se(interaction)"])
)

variable_effects <- data.frame(
  Variable = rownames(migrant_oaxaca$threefold$variables),
  Endowments = migrant_oaxaca$threefold$variables[, "coef(endowments)"],
  Endowments_SE = migrant_oaxaca$threefold$variables[, "se(endowments)"],
  Coefficients = migrant_oaxaca$threefold$variables[, "coef(coefficients)"],
  Coefficients_SE = migrant_oaxaca$threefold$variables[, "se(coefficients)"]
)

selected_variables <- c("AGEP", "SEX", "Citizen", "as.factor(ENG)2", "as.factor(ENG)3",
                        "as.factor(ENG)4", "DIS", "Unemployed", "High_School_Grad",
                        "Bachelors_Degree", "HISP_dummy", "PAP", "HINS4",
                        "SEX:Citizen", "SEX:DIS")

filtered_variable_effects <- variable_effects[variable_effects$Variable %in% selected_variables, ]

ggplot(filtered_variable_effects, aes(x = Variable, y = Endowments, fill = "Endowments")) +
  geom_col(color = "black", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = Endowments - Endowments_SE, ymax = Endowments + Endowments_SE), width = 0.2, position = position_dodge(width = 0.9)) +
  labs(title = "Endowment Effects for Selected Variables  - immigrants", x = NULL, y = "Endowment Effect") +
  scale_fill_manual(values = c("Endowments" = "white")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(filtered_variable_effects, aes(x = Variable, y = Coefficients, fill = "Coefficients")) +
  geom_col(color = "black", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = Coefficients - Coefficients_SE, ymax = Coefficients + Coefficients_SE), width = 0.2, position = position_dodge(width = 0.9)) +
  labs(title = "Coefficient Effects for Selected Variables - immigrants", x = NULL, y = "Coefficient Effect") +
  scale_fill_manual(values = c("Coefficients" = "white")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
coefficients_A <- summary(migrant_oaxaca$reg$reg.A)$coefficients
coefficients_A <- coefficients_A[!grepl("^as\\.factor\\(ST\\)|\\(Intercept\\)", rownames(coefficients_A)), ]
coefficients_A_df <- data.frame(
  Variable = rownames(coefficients_A),
  Coefficient = coefficients_A[, "Estimate"],
  StdError = coefficients_A[, "Std. Error"],
  Model = "Regression immigrants"
)

coefficients_B <- summary(migrant_oaxaca$reg$reg.B)$coefficients
coefficients_B <- coefficients_B[!grepl("^as\\.factor\\(ST\\)|\\(Intercept\\)", rownames(coefficients_B)), ]
coefficients_B_df <- data.frame(
  Variable = rownames(coefficients_B),
  Coefficient = coefficients_B[, "Estimate"],
  StdError = coefficients_B[, "Std. Error"],
  Model = "Regression native-born"
)

all_coefficients <- rbind(coefficients_A_df, coefficients_B_df)

ggplot(all_coefficients, aes(x = Variable, y = Coefficient, fill = Model)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = Coefficient - StdError, ymax = Coefficient + StdError), 
                position = position_dodge(width = 0.8), width = 0.25) +
  coord_flip() +
  scale_fill_manual(values = c("grey70", "grey30")) +
  labs(title = "Comparison of Coefficients",
       x = "Variables",
       y = "Coefficients") +
  theme_minimal() +
  theme(legend.position = "top")
```




EDA
```{r}
library(ggplot2)
library(gridExtra)
library(scales)  # Ensure scales is loaded for percent formatting

generate_and_save_plots <- function(data, variable, file_name) {
  # Base plot with aes_string for dynamic variable use
  base_plot <- ggplot(data, aes_string(x = variable, y = "..prop..", group = "1")) +
    geom_bar(aes(y = ..prop..), stat = "count", fill = "gray") +
    scale_y_continuous(labels = percent_format()) +
    geom_text(stat = "count", aes(label = percent(..prop..), y = ..prop..), 
              vjust = -0.5, position = position_stack(vjust = 0.5))

  # General distribution plot
  p1 <- base_plot +
    labs(title = paste("General Distribution of", variable), x = variable, y = "Proportion")
  
  # Distribution by citizen status
  p2 <- base_plot +
    facet_grid(. ~ Citizen) +
    labs(title = paste("Distribution of", variable, "by Citizen Status"), x = variable, y = "Proportion")
  
  # Distribution by immigrant status
  p3 <- base_plot +
    facet_grid(. ~ immigrant) +
    labs(title = paste("Distribution of", variable, "by Immigrant Status"), x = variable, y = "Proportion")

  # Distribution by Asian dummy status
  p4 <- base_plot +
    facet_grid(. ~ asian_dummy) +
    labs(title = paste("Distribution of", variable, "by Asian identification"), x = variable, y = "Proportion")

  # Distribution by Hispanic dummy status
  p5 <- base_plot +
    facet_grid(. ~ HISP_dummy) +
    labs(title = paste("Distribution of", variable, "by Hispanic identification"), x = variable, y = "Proportion")
  
  # Combine plots vertically
  combined_plot <- grid.arrange(p1, p2, p3, p4, p5, ncol = 1)
  
  # Save the combined plot to a file
  ggsave(file_name, combined_plot, width = 8, height = 30)  # Adjusted height to accommodate more plots
}

# Generate and save plots for each variable
generate_and_save_plots(processed_data2, "PAP_binary", "plots_PAP_binary.png")
generate_and_save_plots(processed_data2, "HINS4", "plots_HINS4.png")
generate_and_save_plots(processed_data2, "Unemployed", "plots_Unemployed.png")
```
