---
title: "230A - Group Project"
format: pdf
editor: source
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(MASS)
```

# Reading in

```{r}
zip_file_path <- "../processed_data.zip"
csv_file_inside_zip <- "processed_data.csv"
processed_data <- read.csv(unz(zip_file_path, csv_file_inside_zip), header = TRUE,
                 sep = ",") 
```

```{r}
processed_data <- processed_data %>%
  mutate(
    DIS = ifelse(DIS == 2, 0, 1),
    HINS4 = 1 - as.numeric(HINS4))
```

```{r}
processed_data$ENG[is.na(processed_data$ENG)] <- 0
```

# Data Dictionary:

ST: State

DIVISION: Division code based on 2010 Census definitions Division code based on 2020 Census definitions 
0 .Puerto Rico 
1 .New England (Northeast region) 
2 .Middle Atlantic (Northeast region) 
3 .East North Central (Midwest region) 
4 .West North Central (Midwest region) 
5 .South Atlantic (South region) 
6 .East South Central (South region) 
7 .West South Central (South Region) 
8 .Mountain (West region)

AGEP: Age 
0 .Under 1 year 
1..99 .1 to 99 years (Top-coded)

ESR: 
1 .Civilian employed, at work 
2 .Civilian employed, with a job but not at work 
3 .Unemployed 
4 .Armed forces, at work 
5 .Armed forces, with a job but not at work

NATIVITY: Nativity 
1 .Native 
2 .Foreign born

CIT: Citizenship status 
1 .Born in the United States 
2 .Born in Puerto Rico, Guam, the U.S. Virgin Islands, or .Northern Marianas 
3 .Born abroad of U.S. citizen parent or parents 
4 .U.S. citizen by naturalization 
5 .Not a U.S. citizen

CITWP: Year of naturalization 
bbbb: N/A (Not eligible - not naturalized)

YOEP: Year of entry 
bbbb .N/A (Not eligible - Born in the US)

COW: Class of worker 
b .N/A (NILF who last worked more than 5 years ago or never worked) 
1 .Employee of a private for-profit company or business, or of an .individual, for wages, salary, or commissions 
2 .Employee of a private not-for-profit, tax-exempt, or .charitable organization 
3 .Local government employee (city, county, etc.) 
4 .State government employee 5 .Federal government employee 
6 .Self-employed in own not incorporated business, professional .practice, or farm 
7 .Self-employed in own incorporated business, professional .practice or farm 
8 .Working without pay in family business or farm 
9 .Unemployed and last worked 5 years ago or earlier or never .worked

Disability: 
1 .With a disability 
2 .Without a disability

ENG: Ability to speak English 
0 .speaks only English 
1 .Very well 
2 .Well 
3 .Not well 
4 .Not at all

PAP: Public assistance income past 12 months (use ADJINC to adjust to constant dollars) 
0 .None 
4..30000 .
\$4 to \$30000 (Rounded and top-coded)

FER: Gave birth to child within the past 12 months 
b .N/A (greater than 50 years/ male) 
1 .Yes 
2. No

MAR: Marital status 
1 .Married 
2 .Widowed 
3 .Divorced 
4 .Separated 
5 .Never married

SEX: Sex 
0 .Male 
1 .Female

SCH: School enrollment 
1 .No, has not attended in the last 3 months 
2 .Yes, public school or public college 
3 .Yes, private school or college or home school

SCHL: Educational attainment 
01 .No schooling completed 
02 .Nursery school, preschool 
03 .Kindergarten 
04 .Grade 1 
05 .Grade 2 
06 .Grade 3 
07 .Grade 4
08 .Grade 5
09 .Grade 6 
10 .Grade 7 
11 .Grade 8 
12 .Grade 9 
13 .Grade 10 
14 .Grade 11 
15 .12th grade - no diploma 
16 .Regular high school diploma 
17 .GED or alternative credential
18 .Some college, but less than 1 year 
19 .1 or more years of college credit, no degree
20 .Associate's degree 
21 .Bachelor's degree
22 .Master's degree 
23 .Professional degree beyond a bachelor's degree 
24 .Doctorate degree

PINCP: Total person's income (use ADJINC to adjust to constant dollars)
0 .None 
-19998 .Loss of \$19998 or more (Rounded and bottom-coded components)
-19997..-1 .Loss \$1 to \$19997 (Rounded components)
1..4209995 .\$1 to \$4209995 (Rounded and top-codedcomponents)

WAGP: Wages or salary income past 12 months (use ADJINC to adjust WAGP to constant dollars)
0 .None 
4..999999 .
\$4 to 999999 (Rounded and top-coded)

ANC: Ancestry recode
1 .Single
2 .Multiple
3 .Unclassified
4 .Not reported 
ANC1P: Recoded Detailed Ancestry - first entry 
ANC2P: Recoded Detailed Ancestry - second entry
POBP: Place of birth (Recode)

RAC1P: Recoded detailed race code 
1 .White alone 
2 .Black or African American alone
3 .American Indian alone 
4 .Alaska Native alone 
5 .American Indian and Alaska Native tribes specified; or American Indian or Alaska Native, not specified and no other races
6 .Asian alone 
7 .Native Hawaiian and Other Pacific Islander alone 
8 .Some Other Race alone 
9 .Two or More Races

PERNP: Total person's earnings (use ADJINC to adjust to constant dollars) 
0 .No earnings 
-10000 .Loss of \$10000 or more (Rounded and bottom-coded components) 
-9999..-1 .Loss \$1 to \$9999 (Rounded components) 
1..1999998 .\$1 to \$1999998 (Rounded and top-coded components)

HISP: Recoded detailed Hispanic origin

HINS4:Medicaid, Medical Assistance, or any kind of government-assistance plan for those with low incomes or a disability 
1 .Yes 
0 .No

Citizen: ifelse(CIT %in% c(1, 2, 3, 4), 1, 0) 
Foreign_born: ifelse(NATIVITY == 2, 1, 0)
High_School_Grad: ifelse(SCHL >= 16 & SCHL < 21, 1, 0) 
Bachelors_Degree: ifelse(SCHL >= 21, 1, 0) 
Unemployed: ifelse(ESR == 3, 1, 0) 
HISP_dummy: ifelse(HISP == 1, 0, 1)
Years_in_US: 2022 - YOEP 
Years_naturalized: 2022 - CITWP Asian: ifelse(RAC1P ==6, 1, 0)
wb50birth: 1: Yes , 0: No/greater than 50 years/ male) 

```{r}
colnames(processed_data)
```
# Cleaning
```{r}
processed_data$Years_naturalized[processed_data$Foreign_born == 0] <- 0
processed_data$Years_in_US[processed_data$Foreign_born == 0] <- 0
processed_data$Years_naturalized[processed_data$CIT == 5] <- 0
processed_data = processed_data%>%filter(CIT != 2)
processed_data$wb50birth <- processed_data$FER
processed_data$wb50birth[is.na(processed_data$FER) &
                         (processed_data$AGEP >= 50 | processed_data$SEX == 0)] <- 0
```

```{r}
#colSums(is.na(processed_data))
```

```{r}
processed_data2 = processed_data%>%filter(WAGP>0 & WAGP < 999999)
processed_data2$adj_WAGP = log(processed_data2$WAGP)
```

```{r}
# Using the boxcox function to determine the best lambda
bc_outcome <- boxcox(WAGP ~ 1, data = processed_data2, lambda = seq(-6, 6, by = 0.1))

# Plotting to visually inspect the best lambda
plot(bc_outcome)

# Finding the lambda that maximizes the log-likelihood
best_lambda <- bc_outcome$x[which.max(bc_outcome$y)]
print(best_lambda)
```


```{r}
# Applying the Box-Cox transformation
processed_data2$bc_WAGP <- if(best_lambda != 0) {
  (processed_data2$WAGP^best_lambda - 1) / best_lambda
} else {
  log(processed_data2$WAGP)
}
```

# EDA Visualizations

```{r}
a <- ggplot(processed_data2,aes(x = WAGP)) + geom_histogram()
b <- ggplot(processed_data2,aes(x = adj_WAGP)) + geom_histogram()
grid.arrange(a,b, ncol = 2)
```

```{r}
wagecit<- ggplot(processed_data2, aes(x = WAGP, fill = factor(Citizen), color = factor(Citizen))) +
  geom_density(aes(y = ..density..), alpha = 0.5) +
  labs(title = "Distribution of Wages by Race",
       x = "Wages",
       y = "Density (Proportion)") + theme_minimal()
wagecit
```

```{r}
logwagecit<- ggplot(processed_data2, aes(x = adj_WAGP, fill = factor(Citizen), color = factor(Citizen))) +
  geom_density(aes(y = ..density..), alpha = 0.5) +
  labs(title = "Distribution of log(Wages + 1) by Race",
       x = "log(Wages + 1)",
       y = "Density (Proportion)") + theme_minimal()
```

```{r}
wagerace<- ggplot(processed_data2, aes(x = WAGP, fill = factor(RAC1P), color = factor(RAC1P))) +
  geom_density(aes(y = ..density..), alpha = 0.5) +
  facet_wrap(vars(factor(Citizen))) +
  labs(title = "Distribution of Wages by Race, Faceted by Citizenship Status",
       x = "Wages",
       y = "Density (Proportion)") + theme_minimal()
```
```{r}
logwagerace<- processed_data2%>%filter(RAC1P %in% c(1,2,6))%>%ggplot(aes(x = adj_WAGP, fill = factor(RAC1P), color = factor(RAC1P))) +
  geom_density(aes(y = ..density..), alpha = 0.5) +
  facet_wrap(vars(factor(Citizen))) +
  labs(title = "Distribution of log(Wages + 1) by Race, Faceted by Citizenship Status",
       x = "log(Wages + 1)",
       y = "Density (Proportion)") + theme_minimal()
```
```{r}
logwage_for<- processed_data2%>%filter(RAC1P %in% c(1,2,6,8,9))%>%ggplot(aes(x = adj_WAGP)) +
  geom_density(aes(y = ..density.., fill = factor(RAC1P), color = factor(RAC1P)), alpha = 0.5) +
  facet_wrap(vars(factor(Foreign_born))) +
  labs(title = "Distribution of log(Wages + 1) by Race, Faceted by Foreign_born",
       x = "log(Wages + 1)",
       y = "Density (Proportion)") + theme_minimal()
```

# Multiple Regression Models

## Predicting Wages in America: 

```{r}
processed_data2 = processed_data2%>%mutate(immigrant = CIT%in%c(4,5))
```


### model0 and summary
```{r}
model0 = lm(adj_WAGP ~as.factor(RAC1P) + AGEP + I(AGEP^2) + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + PAP + HINS4 + immigrant, data = processed_data2)
```

```{r}
# Get the summary of the model
summary_model <- summary(model0)

# Filter out rows related to state coefficients
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]

# Print the refined summary
summary_model
```
### qqplot
```{r}
residuals_df <- data.frame(residuals = resid(model0))
ggplot(residuals_df, aes(sample = residuals)) +
  stat_qq() +
  stat_qq_line(color = "red") +
  ggtitle("Normal Q-Q Plot of Residuals") +
  xlab("Theoretical Quantiles") +
  ylab("Sample Quantiles") +
  theme_minimal() +
  theme(aspect.ratio = 1)
```


```{r}
qqnorm(resid(model0))
qqline(resid(model0))
```
### resid plot

```{r}
# produce residual vs. fitted plot 
plot(fitted(model0), resid(model0)) 
  
# add a horizontal line at 0  
abline(0,0) 
```

### model0.1 and summary

```{r}
model0.1 = lm(bc_WAGP ~as.factor(RAC1P) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + PAP + HINS4 + immigrant, data = processed_data2)
```

```{r}
# Get the summary of the model
summary_model <- summary(model0.1)

# Filter out rows related to state coefficients
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]

# Print the refined summary
summary_model
```
### qqplot
```{r}
qqnorm(resid(model0.1))
qqline(resid(model0.1))
```
### resid plot

```{r}
# produce residual vs. fitted plot 
plot(fitted(model0.1), resid(model0.1)) 
  
# add a horizontal line at 0  
abline(0,0) 
```


## Predicting Immigrants Wages: Differences across Wages for immigrants of different Races


```{r}
immigrants_only = processed_data2%>%filter(CIT%in% c(4,5))
immigrants_only2 = immigrants_only%>%filter(RAC1P%in%c(1,2,6,8,9))
```
### model and summary
```{r}
model1 = lm(adj_WAGP ~ as.factor(RAC1P) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + as.factor(MAR) + PAP + HINS4, data = immigrants_only2)
```

```{r}
# Get the summary of the model
summary_model <- summary(model1)

# Filter out rows related to state coefficients
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]

# Print the refined summary
summary_model
```
### qqplot


```{r}
qqnorm(resid(model1))
qqline(resid(model1))
```


## Predicting Wages for Asian: Differences across Wages for Asian immigrants and from top 5 countries and also non-immigrants


### selecting new data for asians
```{r}
asian_only = processed_data2%>%filter(Asian==1)
asian_only_im = immigrants_only2%>%filter(Asian==1)
```
```{r}
largest_subgroup = asian_only_im%>%group_by(POBP)%>%
  summarise(count = n())%>%
  mutate(per = (count/sum(count))*100)%>%
  arrange(desc(per))%>%head(5)%>%pull(POBP)
```

```{r}
asian_only_im = asian_only_im%>%filter(POBP %in% largest_subgroup)
```

### plots for asians

```{r}
# Create density plot for RAC1P groups
density_plot_rac1p <- ggplot(immigrants_only2 %>% filter(RAC1P %in% c(1, 2, 6)), aes(x = adj_WAGP, fill = factor(RAC1P), color = factor(RAC1P))) +
  geom_density(aes(y = ..density..), alpha = 0.5) +
  labs(title = "Distribution of Immigrant Wages by Race",
       x = "Wages",
       y = "Density (Proportion)") + ylim(0,0.6)

# Create density plot for RAC1P groups
density_plot_rac1p2 <- ggplot(processed_data2 %>% filter(RAC1P %in% c(1, 2, 6) & immigrant ==0), aes(x = adj_WAGP, fill = factor(RAC1P), color = factor(RAC1P))) +
  geom_density(aes(y = ..density..), alpha = 0.5) +
  labs(title = "Distribution of non-Immigrant Wages by Race",
       x = "Wages",
       y = "Density (Proportion)") + ylim(0,0.6)

# Create density plot for Asian immigrants
density_plot_asian <- ggplot(asian_only_im, aes(x = adj_WAGP, fill = factor(POBP), color = factor(POBP))) +
  geom_density(aes(y = ..density..), alpha = 0.5) +
  labs(title = "Distribution of Immigrant Wages for Asians by Country of Origin",
       x = "Wages",
       y = "Density (Proportion)")+ ylim(0,0.6)

density_plot_asian2 <- ggplot(asian_only, aes(x = adj_WAGP)) +
  geom_density(aes(y = ..density..), alpha = 0.5, alpha = 0.5, fill = "cornflowerblue", color = "cornflowerblue") +
  labs(title = "Distribution of non-Immigrants Wages for Asians",
       x = "Wages",
       y = "Density (Proportion)")+ ylim(0,0.6)

# Combine plots side by side
library(gridExtra)
grid.arrange(density_plot_rac1p, density_plot_rac1p2, density_plot_asian, density_plot_asian2, ncol = 2, nrow = 2)

```

### model 1 for asian immigrants

```{r}
model2 = lm(adj_WAGP ~ AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + PAP + HINS4 + as.factor(MAR) , data = asian_only_im)
```

```{r}
# Get the summary of the model
summary_model <- summary(model2)

# Filter out rows related to state coefficients
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]

# Print the refined summary
summary_model
```
### qqplot1
```{r}
qqnorm(resid(model2))
qqline(resid(model2))
```
### model 2 for all asians

```{r}
model2.1 = lm(adj_WAGP ~ AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth + Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + HISP_dummy + PAP + HINS4 + immigrant + as.factor(MAR) , data = asian_only)
```

```{r}
# Get the summary of the model
summary_model <- summary(model2.1)

# Filter out rows related to state coefficients
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]

# Print the refined summary
summary_model
```

### qqplot2
```{r}
qqnorm(resid(model2.1))
qqline(resid(model2.1))
```



## Predicting Immigrants Wages: Differences across Wages for Hispanic immigrants from top 5 countries and also non-immigrants

### HISP data cleaning
```{r}
hisp_only = processed_data2%>%filter(HISP_dummy==1)
hisp_only_im = immigrants_only2%>%filter(HISP_dummy==1)
```
```{r}
largest_subgroup = hisp_only_im%>%group_by(POBP)%>%
  summarise(count = n())%>%
  mutate(per = (count/sum(count))*100)%>%
  arrange(desc(per))%>%head(5)%>%pull(POBP)
```

```{r}
hisp_only_im = hisp_only_im%>%filter(POBP %in% largest_subgroup)
```

### HISP EDA

```{r}
# Create density plot for RAC1P groups
density_plot_rac1p <- ggplot(hisp_only_im, aes(x = adj_WAGP, fill = factor(RAC1P), color = factor(RAC1P))) +
  geom_density(aes(y = ..density..), alpha = 0.5) +
  labs(title = "Distribution of Hispanic Immigrant Wages by Race",
       x = "Wages",
       y = "Density (Proportion)") + ylim(0,0.6)

# Create density plot for RAC1P groups
density_plot_rac1p2 <- ggplot(hisp_only %>% filter(immigrant ==0), aes(x = adj_WAGP, fill = factor(RAC1P), color = factor(RAC1P))) +
  geom_density(aes(y = ..density..), alpha = 0.5) +
  labs(title = "Distribution of non-Immigrant Hispanic Wages by Race",
       x = "Wages",
       y = "Density (Proportion)") + ylim(0,0.6)

# Create density plot for Hispanic immigrants
density_plot_hisp <- ggplot(hisp_only_im, aes(x = adj_WAGP, fill = factor(POBP), color = factor(POBP))) +
  geom_density(aes(y = ..density..), alpha = 0.5) +
  labs(title = "Distribution of Immigrant Wages for Hispanics by Country of Origin",
       x = "Wages",
       y = "Density (Proportion)")+ ylim(0,0.6)

density_plot_hisp2 <- ggplot(hisp_only%>% filter(immigrant ==0), aes(x = adj_WAGP)) +
  geom_density(aes(y = ..density..), alpha = 0.5, alpha = 0.5, fill = "cornflowerblue", color = "cornflowerblue") +
  labs(title = "Distribution of non-Immigrant Wages for Hispanics",
       x = "Wages",
       y = "Density (Proportion)")+ ylim(0,0.6)

density_plot_hisp3 <- ggplot(hisp_only_im, aes(x = adj_WAGP)) +
  geom_density(aes(y = ..density..), alpha = 0.5, alpha = 0.5, fill = "cornflowerblue", color = "cornflowerblue") +
  labs(title = "Distribution of Immigrant Wages for Hispanics",
       x = "Wages",
       y = "Density (Proportion)")+ ylim(0,0.6)

# Combine plots side by side
library(gridExtra)
grid.arrange(density_plot_rac1p, density_plot_rac1p2, density_plot_hisp, density_plot_hisp2,density_plot_hisp3, ncol = 2, nrow = 3)
```

### HISP model 1 for immigrants

```{r}
model3 = lm(adj_WAGP ~ as.factor(RAC1P) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + PAP + HINS4 + as.factor(MAR), data = hisp_only_im)
```

```{r}
# Get the summary of the model
summary_model <- summary(model3)

# Filter out rows related to state coefficients
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]

# Print the refined summary
summary_model
```

### qqplot 1
```{r}
qqnorm(resid(model3))
qqline(resid(model3))
```

### HISP model 2 for all hispanics

```{r}
model3.1 = lm(adj_WAGP ~ as.factor(RAC1P) + AGEP + SEX*Citizen + as.factor(ENG) + SEX*DIS + wb50birth+ Unemployed + as.factor(ST) + High_School_Grad + Bachelors_Degree + PAP + HINS4 + immigrant + as.factor(MAR), data = hisp_only)
```

```{r}
# Get the summary of the model
summary_model <- summary(model3.1)

# Filter out rows related to state coefficients
state_rows <- grepl("^as\\.factor\\(ST\\)", rownames(summary_model$coefficients))
summary_model$coefficients <- summary_model$coefficients[!state_rows, ]

# Print the refined summary
summary_model
```
### qqplot2

```{r}
qqnorm(resid(model3.1))
qqline(resid(model3.1))
```


